---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


```{r setup, include=TRUE, echo = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(DT)
```

## Loading and preprocessing the data

```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",
              destfile = "./Data/activity.zip")
unzip("./Data/activity.zip", exdir = "./Data")
data <- read.csv("./Data/activity.csv") %>%
    mutate(date = as.Date(date))
```
##

## What is mean total number of steps taken per day?

### Steps Taken by Day

```{r}
# Calculate steps taken by day
summary <- data %>%
    group_by(date) %>%
    summarise(totalSteps = sum(steps))
datatable(summary)
```

### Distribution of Steps Taken by Day

```{r}
ggplot(data = summary, aes(totalSteps)) +
    geom_histogram()
```

The mean of steps in a day is `r mean(summary$totalSteps, na.rm = TRUE)`.
The median is `r median(summary$totalSteps, na.rm = TRUE)`.

##

## What is the average daily activity pattern?

### Average Daily Activity per interval

```{r}
summary <- data %>%
    group_by(interval) %>%
    summarise(stepsMean = mean(steps, na.rm = TRUE))
datatable(summary)
```

### Daily Activity Distribution

```{r}
ggplot(data = summary) +
    geom_line(aes(interval, stepsMean))
```

The interval that contains the maximum number of steps mean is
`r summary %>% with(interval[stepsMean == max(stepsMean)])`.

##

## Imputing missing values

### Missing Values Countting

There are `sum(is.na(summary$steps))` missing values n the set.

### Filling Missing Values

The missing values are going to be replaced by the mean of its interval across the days.

```{r}
data2 = data %>% 
  merge(summary) %>%
  (function(X) {
    nas = is.na(X$steps)
    X$steps[nas] = round(X$stepsMean[nas], 2)
    return(X)
  }) %>%
  select(3:1)
datatable(data2)
```

### New Distribution of Steps Taken by Day

```{r}
# Calculate steps taken by day
summary2 <- data2 %>%
    group_by(date) %>%
    summarise(totalSteps = sum(steps))
datatable(summary2)
```

### Distribution of Steps Taken by Day

```{r}
ggplot(data = summary2, aes(totalSteps)) +
    geom_histogram()
```

The mean of steps in a day is `r mean(summary2$totalSteps, na.rm = TRUE)`.
The median is `r median(summary2$totalSteps, na.rm = TRUE)`.

##

## Are there differences in activity patterns between weekdays and weekends?

```{r}
data2 %>% 
  mutate(day = factor(weekdays(data2$date) %in% c("sábado", "domingo"), 
                      labels = c("weekday", "weekend"))) %>%
  group_by(interval, day) %>%
  summarise(stepsMean = mean(steps, na.rm = T)) %>%
  ggplot(aes(interval, stepsMean)) +
  geom_line() +
  facet_wrap(vars(day), ncol = 1)
```











